Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-05-05 paper.tex
Remove for saving space

Interestingly, many of non-volatile memory technologies are making their ways
into storage drives. One potential use of fast storage devices is for fast
swapping space. In such a system, the overhead of swapping may become a
bottleneck.

Prefetching pages from swap space can offer a way to mitigate the overhead of
memory overcommitment, possibly making large virtual machines a viable
mechanism for studying large memory systems. When there is significant memory
pressure, even pages which are likely to be accessed soon are swapped out to
disk. They are faulted back into memory when accessed, resulting in significant
performance cost.  Charm++ uses a programming model where computations can be
scheduled by the language runtime. A designated thread can prefetch pages
required by a computation, averting page faults \cite{charmpp}. Rather than
implement page prefetching in a language runtime, one might implement a
prefetcher in the swapping subsystem, making it general purpose. A large body
of work already exists on hardware prefetching for processor caches, on which a
swap prefetcher might draw for inspiration \cite{prefetching}. However, this
approach can only work if page access patterns are predictable. Moreover, the
high latency of disks currently implies that prefetchers would have to predict
the very distant future (e.g. seconds ahead of execution); however, with
non-volatile memory-based swap space, this idea may be more practical.

@@

(TODO: where to put this paragraph?)
A study of recent changes to memory management in Linux has found that there
has been a lot of focus on improving the scalability of data structures used in
the Linux memory manager \cite{huang2016evolutionary}. Since systems with
persistent memory are expected to have large amounts of memory, there have been
efforts to reduce the overheads of \texttt{struct page}. This is because these
structures can consume several gigabytes of memory for systems with terabytes
of persistent memory \cite{corbet_persistent_progress}. These structures need
to be stored in volatile memory not only because they need to be accessed
quickly for performance reasons, but also to avoid wearing out of persistent
memory due to frequent writes to this data structure. Patches proposed to the
Linux kernel have proposed getting rid of the \texttt{struct page} altogether, but the
complexity associated with such radical changes have hindered their acceptance
by the community. Other works such as RadixVM \cite{radixvm} have also
approached scalability with respect to the number of cores in the system.

>>>>

%Interestingly, many of non-volatile memory technologies are making their ways
%into storage drives. One potential use of fast storage devices is for fast
%swapping space. In such a system, the overhead of swapping may become a
%bottleneck.
%
%Prefetching pages from swap space can offer a way to mitigate the overhead of
%memory overcommitment, possibly making large virtual machines a viable
%mechanism for studying large memory systems. When there is significant memory
%pressure, even pages which are likely to be accessed soon are swapped out to
%disk. They are faulted back into memory when accessed, resulting in significant
%performance cost.  Charm++ uses a programming model where computations can be
%scheduled by the language runtime. A designated thread can prefetch pages
%required by a computation, averting page faults \cite{charmpp}. Rather than
%implement page prefetching in a language runtime, one might implement a
%prefetcher in the swapping subsystem, making it general purpose. A large body
%of work already exists on hardware prefetching for processor caches, on which a
%swap prefetcher might draw for inspiration \cite{prefetching}. However, this
%approach can only work if page access patterns are predictable. Moreover, the
%high latency of disks currently implies that prefetchers would have to predict
%the very distant future (e.g. seconds ahead of execution); however, with
%non-volatile memory-based swap space, this idea may be more practical.

@@

(TODO: where to put this paragraph?)
A study of recent changes to memory management in Linux has found that there
has been a lot of focus on improving the scalability of data structures used in
the Linux memory manager \cite{huang2016evolutionary}. Since systems with
persistent memory are expected to have large amounts of memory, there have been
efforts to reduce the overheads of \texttt{struct page}. This is because these
structures can consume several gigabytes of memory for systems with terabytes
of persistent memory \cite{corbet_persistent_progress}. These structures need
to be stored in volatile memory not only because they need to be accessed
quickly for performance reasons, but also to avoid wearing out of persistent
memory due to frequent writes to this data structure. Patches proposed to the
Linux kernel have proposed getting rid of the \texttt{struct page} altogether, but the
complexity associated with such radical changes have hindered their acceptance
by the community.
